---
title: "Intro to DATA662 and mldash Package"
author: "Angela Lui, Ph.D. and Jason Bryer, Ph.D."
format: 
  revealjs:
    theme: [simple]
    preview-links: auto
    chalkboard: 
      boardmarker-width: 5
execute:
  echo: true
editor: source
---

## Welcome to DATA 661!

-   Instructors: [Angela Lui, Ph.D.](mailto:angela.lui@cuny.edu) and [Jason Bryer, Ph.D.](mailto:jason.bryer@cuny.edu).

-   Course website (including syllabus and schedule): https://jbryer.github.io/DATA661-2022-Fall/

-   `mldash` Package Website: https://github.com/jbryer/mldash

## Getting Started (prerequisites)

Install `reticulate`, `tensorflow`, `keras`, and `torch` from Github. Be warned this will take a while!

```{r}
#| eval: false
remotes::install_github(sprintf("rstudio/%s", c("reticulate", "tensorflow", "keras", "torch")))
reticulate::miniconda_uninstall()
reticulate::install_miniconda()
reticulate::conda_create("mldash")
reticulate::conda_install("mldash",
	c("jupyterlab", "pandas", "statsmodels","scipy", "scikit-learn",
	  "matplotlib", "seaborn", "numpy", "pytorch", "tensorflow"))
install.packages("BiocManager")
BiocManager::install("mixOmics")
```

## Installing `mldash`

```{r}
#| eval: false
remotes::install_github('jbryer/mldash')
```

Some Python configuration...

```{r}
#| warning: false
#| eval: false
Sys.setenv("RETICULATE_PYTHON" = "~/miniforge3/envs/mldash/bin/python")
reticulate::use_condaenv("mldash")
tensorflow::use_condaenv("mldash")
keras::use_condaenv("mldash")
```

Now we can use it 🤞.

```{r}
#| eval: false
library(mldash)
```

## Datasets

```{r}
#| eval: false
datasets <- mldash::read_ml_datasets()
datasets
```

## Models

This will get all the models currently in the package. This may take a while the first time as it will try to install any dependencies defined.

```{r}
#| eval: false
models <- mldash::read_ml_models()
```

## Defining a Dataset

```{r}
#| eval: false
adult_data <- mldash::new_dataset(
	name = 'adult',
	type = 'classification',
	description = 'Prediction task is to determine whether a person makes over 50K a year.',
	source = 'https://archive.ics.uci.edu/ml/datasets/Adult',
	url = 'https://archive.ics.uci.edu/ml/machine-learning-databases/adult/adult.data',
	dir = 'inst/datasets',
	data = function() {
		destfile <- tempfile()
		download.file("https://archive.ics.uci.edu/ml/machine-learning-databases/adult/adult.data", destfile)
		df <- read.csv(destfile, header = FALSE)
		names(df) <- c('age', 'workclass', 'fnlwgt', 'education', 'education_num', 'marital_status',
					   'occupation', 'relationship', 'race', 'sex', 'capital_gain', 'captial_loss',
					   'hours_per_week', 'native_country', 'greater_than_50k')
		df$greater_than_50k <- df$greater_than_50k == ' >50K'
		return(df)
	},
	model = greater_than_50k ~ .,
	overwrite = TRUE
)
```

## Defining a Dataset (result)

```
name: adult
type: classification
description: Prediction task is to determine whether a person makes over 50K a year.
source: https://archive.ics.uci.edu/ml/datasets/Adult
reference: APA reference for the dataset.
url: https://archive.ics.uci.edu/ml/machine-learning-databases/adult/adult.data
data: function () 
    {
        destfile <- tempfile()
        download.file("https://archive.ics.uci.edu/ml/machine-learning-databases/adult/adult.data", 
            destfile)
        df <- read.csv(destfile, header = FALSE)
        names(df) <- c("age", "workclass", "fnlwgt", "education", 
            "education-num", "marital-status", "occupation", "relationship", 
            "race", "sex", "capital-gain", "captial-loss", "hours-per-week", 
            "native-country", "greater_than_50k")
        df$greater_than_50k <- df$greater_than_50k == " >50K"
        return(df)
    }
model: greater_than_50k ~ .
note:
```

## Defining a Model

```{r}
#| eval: false
rf_model <- mldash::new_model(
	name = 'randomForest_classification',
	type = 'classification',
	description = 'Random forest prediction model usign the randomForest R package.',
	train_fun = function(formula, data) {
		randomForest::randomForest(formula = formula, data = data, ntree = 1000)
	},
	predict_fun = function(model, newdata) {
	    randomForest:::predict.randomForest(model, newdata = newdata, type = "prob")[,2,drop=TRUE]
	},
	packages = "randomForest",
	overwrite = TRUE
)
```

## Defining a Model (result)

```
name: randomForest_classification
type: classification
description: Random forest prediction model usign the randomForest R package.
train: function (formula, data)
    {
        randomForest::randomForest(formula = formula, data = data,
            ntree = 1000)
    }
predict: function (model, newdata)
    {
        randomForest:::predict.randomForest(model, newdata = newdata, type = "prob")[,2,drop=TRUE]
    }
packages: randomForest
note:
```

## Running Models

```{r}
#| eval: false
ml_results <- mldash::run_models(datasets = datasets, 
								 models = models,
								 print_errors = FALSE, 
								 seed = 1234)
```

## Next Steps

Before next week try to:

1.  Install the `mldash` R package.

2.  Run the models and datasets already there.

3.  Find a dataset from the UCI Machine Learning Repository: https://archive.ics.uci.edu/ml/index.php When you find one you want to work on add it to this [Google Sheet](https://docs.google.com/spreadsheets/d/1z7z5AiDIB9ImBduoLR0-Ycvvtv-nRfAtPbNXAoogIz4/edit?usp=sharing). We will use this to ensure we don't step on someone's toes:
